---
title: "Final Project Data Evaluation"
author: "Anya Weaver"
# format: live-html
# engine: knitr
format: #revealjs
    pdf:
      keep-tex: true
      include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
            \DefineVerbatimEnvironment{OutputCode}{Verbatim}{breaklines,commandchars=\\\{\}}
    html:
      self-contained: true
      grid:
        margin-width: 200px
      code-fold: false
      toc: true
      # callout-appearance: minimal
# You can change the color theme of the rendered document 
theme: Yeti
---

```{r include = FALSE}
knitr::opts_chunk$set(
  error = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.height = 2.75, 
  fig.width = 4.25,
  fig.env='figure',
  fig.pos = 'h',
  fig.align = 'center')
```

---
title: "Data Evaluation for Final Project"
subtitle: "STAT 244"
author: "Anya Weaver"
output:
  pdf_document:
    fig_height: 1.5
    fig_width: 4
    highlight: zenburn
header-includes:
  - \usepackage{booktabs}
---

Setup file

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(readr)
library(purrr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(mosaic)
library(gmodels)
library(Sleuth3)
```

Load dataset of interest:

```{r}
#commented out because I made a new CSV
#field = read_csv("Field.csv")
```

Rename CIPDESC to major for clarity:

```{r}
#field = rename(field, Major = CIPDESC)
```

Asses contents of Dataset:

```{r}
#head(field)
```

Sort by Major to asses options:

```{r}
# commented out b/c it takes a while, makes a lot of data, and I don't need to do it again
#table(field$major)
```

Mutate Dataset to only include Biology, General and  Biochemistry, Biophysics and Molecular Biology Majors & rename those majors to biology and biochem+:

```{r}
#field = field  %>% mutate(Major = ifelse(Major == 'Biochemistry, Biophysics and Molecular Biology.', "biochem+", Major))

#field = field  %>% mutate(Major = ifelse(Major == 'Biology, General.', "biology", Major))

#bio <- field %>% filter((Major == 'biochem+') | (Major == 'biology'))

#bio <- bio %>% mutate(Major = as.factor(Major))
```

Turn new dataset into its own CSV so I don't have to load field again:

```{r}
#commented out because I only need the one CSV
#write_csv(bio, "bio.csv")
#bio = read_csv("bio.csv")
```

Mutate Dataset to only include some variables:

```{r}
#figure out what column names translate to the columns DEBTMEDIAN, DEBTMEAN, and MD_EARN_WNE. Figure out how gender comes into it. 

# bio <- bio %>% select(Major, INSTNM, CONTROL, CREDDESC, DEBT_ALL_PP_EVAL_MDN, DEBT_ALL_STGP_EVAL_MDN, EARN_MDN_HI_1YR, DISTANCE, EARN_MDN_5YR)
# head(bio)
```

Filter out non responses:

```{r}
# bio = bio %>% filter(!is.na(DEBT_ALL_PP_EVAL_MDN) & (DEBT_ALL_PP_EVAL_MDN !="PS"))
```

Add STGPand PP together to make one value for debt:

```{r}
# bio$DEBT_ALL_PP_EVAL_MDN = as.numeric(bio$DEBT_ALL_PP_EVAL_MDN)
# bio$DEBT_ALL_STGP_EVAL_MDN = as.numeric(bio$DEBT_ALL_STGP_EVAL_MDN)
# bio$Debt_median <- bio$DEBT_ALL_PP_EVAL_MDN + bio$DEBT_ALL_STGP_EVAL_MDN
```

Rename included variables for clarity:

```{r}
# bio = rename(bio, Institution = INSTNM)
# bio = rename(bio, School_type = CONTROL)
# bio = rename(bio, Degree = CREDDESC)
# bio = rename(bio, Distance = DISTANCE)
# bio = rename(bio, Earnings_median_1yr = EARN_MDN_HI_1YR)
# bio = rename(bio, Earnings_median_5yr = EARN_MDN_5YR)
```

Remove old columns:

```{r}
# bio <- bio %>% select(Major, Institution, 
#                           School_type, Degree, Debt_median, Distance, Earnings_median_1yr, Earnings_median_5yr)
# head(bio)
```
Filter out NAs:

```{r}
# bio$Earnings_median_1yr = as.numeric(bio$Earnings_median_1yr)
# bio$Earnings_median_5yr = as.numeric(bio$Earnings_median_5yr)
```


Make sure quantitative variables are recorded as numbers:

```{r}
# bio = bio %>% filter(!is.na(Earnings_median_1yr) & (Earnings_median_5yr !="PS"))
# bio$Earnings_median_1yr = as.numeric(bio$Earnings_median_1yr)
# bio$Earnings_median_5yr = as.numeric(bio$Earnings_median_5yr)
```

Read Bio:

```{r}
# write_csv(bio, "bio.csv")
bio = read_csv("bio.csv")
```

Analyze Quantitative Variables:

```{r}
mean(bio$Debt_median)
sd(bio$Debt_median)

mean(bio$Earnings_median_1yr)
sd(bio$Earnings_median_1yr)

mean(bio$Earnings_median_5yr)
sd(bio$Earnings_median_5yr)
```

Analyze Qualitative Variables:

```{r}
count(bio$Institution)

count(bio$School_type)

count(bio$Degree)

count(bio$Distance)

```

Produce data visualizations

```{r}
bio %>%
ggplot(aes(x = Debt_median, y = Earnings_median_5yr)) +
geom_point() +
labs(x = 'Median Debt ($)', y = 'Median Earnings 5 Years After Graduation ($)') +
theme_classic()


bio %>%
ggplot(aes(x = Earnings_median_1yr, y = Earnings_median_5yr)) +
geom_point() +
labs(x = 'Median Earnings 1 Year After Graduation ($)', y = 'Median Earnings 5 Years After Graduation ($)') +
theme_classic()

bio %>%
ggplot(aes(x = Debt_median, y = Earnings_median_1yr)) +
geom_point() +
labs(x = 'Median Debt ($', y = 'Median Earnings 1 Year After Graduation ($)') +
theme_classic()

```

Histograms for dataset:

```{r}

```


