---
title: "STAT 244-SC"
format: html
editor: visual
---

## Final Project

This webpage will have the results of my final project for STAT 244-SC. I am analyzing the link between student debt and median earnings 1 and 5 years after graduation for biochemistry and biology majors according to the College Scorecard project.

## Basic Information

Here are some figures that provide an introduction to the dataset I am analyzing.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(readr)
library(purrr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(mosaic)
library(gmodels)
library(Sleuth3)
bio = read_csv("bio.csv")
```

```{r, echo: false}
bio %>%
ggplot(aes(x = Debt_median, y = Earnings_median_5yr)) +
geom_point() +
labs(x = 'Median Debt ($)', y = 'Median Earnings 5 Years After Graduation ($)') +
theme_classic()


bio %>%
ggplot(aes(x = Earnings_median_1yr, y = Earnings_median_5yr)) +
geom_point() +
labs(x = 'Median Earnings 1 Year After Graduation ($)', y = 'Median Earnings 5 Years After Graduation ($)') +
theme_classic()
```

The `echo: false` option disables the printing of code (only output is displayed).
